Local PCA Shows How Population Structure Differs Along the Genome

### 2016-08-10 writeup/local_pca_paper.tex
.

The phrase ``population structure'',
used to describe patterns in genome-wide mean kinship,
calls to mind to reduced gene flow between subpopulations. %, often because of geographical isolation.
However, it is widely recognized that because of selection the effects of gene flow are not equal everywhere on the genome,
and patterns of polymorphism and divergence can vary significantly depending on factors including local gene density.
This implies that, counterintuitively, 
the population structure of a species depends on which part of the genome is being examined.

@@

Realized kinship summarizes the shapes
of the genealogical trees % that relate the samples
at each location along the genome.
Since these trees vary along the genome, so does realized kinship;
but averaging over sufficiently many trees we hope to get a stable estimate,
that doesn't depend much on the genetic markers chosen.
This is not guarenteed:
for instance,
kinship on sex chromosomes is expected to differ from the autosomes;
and positive or negative selection on particular loci can dramatically disort shapes of nearby genealogies
\citep{maynardsmith1974hitchhiking,barton2000genetic,neher2013genetic,charlesworth2012effects}.
Indeed,
many species show chromosome-scale variation in diversity and divergence
\citep[sometimes dramatic:][]{langley2012genomic};
species phylogenies can differ along the genome 
due to incomplete lineage sorting \citep[e.g.,][]{pease2013accurate},
adaptive introgression and/or local adaptation \citep[e.g.,][]{ellegren2012genomic,nadeau2012genomic,pool2015natural,vernot2014resurrecting};
and theoretical expectations predict that geographic patterns of relatedness should depend on selection
\citep{charlesworth2003review}.
Nonetheless, it is not generally known to what extent patterns of kinship vary along the genome,
nor for what reasons.

@@

\section{Methods}

>>>>

The phrase ``population structure'',
used to describe patterns in genome-wide mean kinship,
calls to mind to reduced gene flow between subpopulations. %, often because of geographical isolation.
However, it is widely recognized that because of selection the effects of gene flow are not equal everywhere on the genome,
and patterns of polymorphism within species and divergence between species can vary significantly depending on factors including local gene density.
This implies that, counterintuitively, 
the population structure of a species depends on which part of the genome is being examined.

@@

Realized kinship summarizes the shapes
of the genealogical trees % that relate the samples
at each location along the genome.
Since these trees vary along the genome, so does realized kinship;
but averaging over sufficiently many trees we hope to get a stable estimate,
that doesn't depend much on the genetic markers chosen.
This is not guarenteed:
for instance,
kinship on sex chromosomes is expected to differ from the autosomes;
and positive or negative selection on particular loci can dramatically disort shapes of nearby genealogies
\citep{kim2002hitchhiking,charlesworth1993effect,barton2000genetic}.
Indeed,
many species show chromosome-scale variation in diversity and divergence
\citep[e.g.,][]{langley2012genomic};
species phylogenies can differ along the genome 
due to incomplete lineage sorting \citep[e.g.,][]{pease2013accurate},
adaptive introgression and/or local adaptation \citep[e.g.,][]{ellegren2012genomic,nadeau2012genomic,pool2015natural,vernot2014resurrecting};
and theoretical expectations predict that geographic patterns of relatedness should depend on selection
\citep{charlesworth2003review}.

@@

Population genetics studies often describe variation in particular statistics along the genome,
often with the aim of detecting particular signals.
In this paper, we use a visualization technique to obtain an unbiased view
of how patterns of kinship vary along the genome,
in genome-wide, spatially distributed datasets for three species,
to start to discover common patterns and differences.

%%%
\section{Methods}
